---
title "Catastrophic Climate Events That Caused US Economic Disorder"
author "Gustavo I. Camilo"
date "10 november 2019"
output html_document
---

```{r}
require(tidyverse)
require(dplyr)
require(pacman)
require(ggplot2)
require(gridExtra)
```

##Loading and Reading Data

```{r}
StormData <- tbl_df(read.csv(file.choose(), header = T, sep = ","))
```

##Data

The data for this assignment come in the form of a comma-separated-value file compressed via the bzip2 algorithm to reduce its size. You can download the file from the course web site:

[DataStorm](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)

There is also some documentation of the database available. Here you will find how some of the variables are constructed/defined.

National Weather Service [Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf)
National Climatic Data Center Storm Events [FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)

The events in the database start in the year 1950 and end in November 2011. In the earlier years of the database there are generally fewer events recorded, most likely due to a lack of good records. More recent years should be considered more complete.


#Data Processing
```{r}
#
StormData.fatalities <- subset(StormData, !StormData$FATALITIES == 0, select = c(EVTYPE, FATALITIES))

#
StormData.injuries <- subset(StormData,!StormData$INJURIES == 
    0, select = c(EVTYPE, INJURIES))

#
EUA_economic <- subset(StormData, !StormData$PROPDMG == 0 & !StormData$CROPDMG == 
    0, select = c(EVTYPE, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP))

```


## 1- Across the United States, which types of events (as indicated in the 'EVTYPE' variable) are most harmful with respect to population health?



```{r}

StormData.fatalities <- aggregate(StormData.fatalities$FATALITIES, by = list(StormData.fatalities$EVTYPE), 
    FUN = sum)
colnames(StormData.fatalities) <- c("EVTYPES", "FATALITIES")
 StormData.fatalities <- StormData.fatalities[order(StormData.fatalities$FATALITIES, decreasing = TRUE), ][1:10, ]
 
 #
 fatalities <- ggplot() + geom_bar(data = StormData.fatalities, aes(x = EVTYPES, 
    y = FATALITIES, fill = interaction(FATALITIES, EVTYPES)), stat = "identity", 
    show.legend = F) + theme(axis.text.x = element_text(angle = 30, hjust = 1)) + 
    xlab("Harmful Events") + ylab("No. of fatailities") + ggtitle("Top 10 weather events causing fatalities") + 
    theme(axis.text.x = element_text(angle = 30, hjust = 1))

```


```{r}


StormData.injuries <- aggregate(StormData.injuries$INJURIES, by = list(StormData.injuries$EVTYPE), FUN = sum)
colnames(StormData.injuries) <- c("EVTYPES","INJURIES")  
StormData.injuries <- StormData.injuries[order(StormData.injuries$INJURIES, decreasing = T), ][1:10, ]

#
injuries <- ggplot() + geom_bar(data = StormData.injuries, aes(x = EVTYPES, y = INJURIES, 
    fill = interaction(INJURIES, EVTYPES)), stat = "identity", show.legend = F) + 
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) + xlab("Harmful Events") + 
    ylab("No. of Injuries") + ggtitle("Top 10 weather events causing Injuries") + 
    theme(axis.text.x = element_text(angle = 30, hjust = 1))


```

#Results
#Top 10 Top Causes of Death and Damage
```{r}

# 
grid.arrange(fatalities, injuries, ncol = 2)

```



##2- Across the United States, which types of events have the #greatest economic consequences?



##Data Processing

```{r}
##Verificando quais limpezas precisamos fazer na coluna 'PROPDMGEXP' indentificado através do levels
head(EUA_economic$PROPDMGEXP)

##Verificando quais limpezas precisamos fazer na coluna 'PROPDMGEXP' indentificado através do levels
head(EUA_economic$CROPDMGEXP)
```
```{r}
##Limpando os dados das tabelas verificadas

EUA_economic <-  EUA_economic %>% mutate_at(vars(PROPDMGEXP,CROPDMGEXP), 
               funs(as.numeric(recode(.,'0'=1,'1'=10,'2'=10^2,'3'=10^3,'4'=10^4,'5'=10^5,
                                             '6'=10^6,'7'=10^7,'8'=10^8, 'B'=10^9,
                                             'h'=1^2,'H'=1^2, 'k'=10^3,'K'=10^3,'m'=10^6,'M'=10^6,
                                             ' '=0,'-'=0,'?'=0,'+'=0))))                            


```


```{r}

PROPDMG <- EUA_economic$PROPDMG * EUA_economic$PROPDMGEXP
CROPDMG <- EUA_economic$CROPDMG * EUA_economic$CROPDMGEXP

EUA_economic$TotalPropCrop <- PROPDMG + CROPDMG

EUA_economic <- aggregate(EUA_economic$TotalPropCrop, by = list(EUA_economic$EVTYPE), FUN = sum) 

colnames(EUA_economic) <- c("EVTYPE", "TotalPropCrop")

EUA_economic <- EUA_economic[order(EUA_economic$TotalPropCrop, decreasing = T), ][1:5,]

head(EUA_economic , n = 10)
```
##Results
```{r}
ggplot() + geom_bar(data = EUA_economic, aes(x = EVTYPE, y = TotalPropCrop, fill = interaction(TotalPropCrop, EVTYPE)), stat = "identity", show.legend = F) + theme(axis.title.x = element_text(angle = 30, hjust = 1)) + ggtitle("Total catastrophic injury and death in the US") + xlab("Event") + ylab("Total ") + coord_flip()

```

#Final Results 

##Based on the results obtained, and with graphs generated, some results that the tornado caused the most deaths and injuries, while the flood caused the economic collapse.
